---
title: "StudyNumber - Title"
author: "FName LName"
date: "11/17/2020"
output: 
  html_document:
    toc: yes
    toc_float: true
    highlight: zenburn
    theme: united
---



1. Chunk Primary Settings:
* There are multiple options to set chunk parameters:
  * Specify global settings to apply to all chunks
```{r setup}
knitr::opts_chunk$set(echo = FALSE, include = FALSE, eval = FALSE, 
                      message = FALSE, warning = FALSE)
print('Name1')
```
  * Create chunk labels that are called later
```{r label = 'chunk-output-only', include=TRUE, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}
print('{r label = 'chunk-output-only', include=TRUE, eval = FALSE, echo = TRUE, message = FALSE, warning = FALSE}')
```
  * Specify on case by case basis through out document
```{r include=TRUE, eval = TRUE, echo = TRUE}
print('{r include=TRUE, eval = TRUE, echo = TRUE}')
```

2. Reporting Standards
* Headings
  * BioMarin Logo present in all outputs
* Body
  * All tables have caption
  * All figures have caption or fig.cap
  * References or link to location required when controlled documents listed
* References
  * All data inputs are listed preferably using fs package 
  * sessionInfo to document package version used in any anlaysis that results in decision with respect to data integrity, safety or study endpoints

  
* Common Packages
  * 'haven', 'stringr', 'dplyr', 'tidyr', 'lubridate', 'purrr', 'fs', 'ggplot2', 'magrittr', 'janitor'





