---
title: "270-301 Covance Topline Metrics"
author: "Jason Delosh"
date: "11/6/2020"
output: 
  html_document:
    toc: yes
    toc_float: true
    highlight: zenburn
    theme: united

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)

library(knitr); library(kableExtra)
```

## **Introduction**

The purpose of the Covance external lab topline metrics report is to identify missing lab results for a subset of laboratory tests specified in the Central Lab Topline Metrics specification document, dated 13Oct2020.  Initial analysis identified a significant number of subjects where one or more analyte results were missing for the specified Covance Lab Analyte.

To maintain all blinding, all laboratory results have been removed from this analysis.

### **Purpose:**

Discussions from the 30Oct2020 meeting determined further exploration is needed to determine if the specification document needed to be updated to more accurately identify tests of interest.  This document outlines the results of that exploration.

### **Inputs:**
* 270-301 EDC SAS on Demand exported data:
    * 'sv.sas7bdat'   =  subject visit data
    * 'ds.sas7bdat'   = subject disposition data
    * 'dsic.sas7bdat' = informed consent data
    * 'dssh.sas7bdat' = previous study history data
    * 'dsss.sas7bdat' =  screening status
    * 'ie.sas7bdat'   =  eligibility data

* External lab data:
  * 'bmn_270301_covance_lab_12oct20.sas7bdat'
  * 'BMN 270-301_bloodworks_mutation_11Feb2020.csv'
  * 'bmn_270301_lb_blnd_fviii_12oct20.sas7bdat'



### **Source Scripts**
'270-301_CovanceTopLineMetrics.R'
```{r, include=FALSE}
source('270-301_CovanceTopLineMetrics.R', local = knitr::knit_global())
```


## **Data Review**

### **Subjects**
This analysis includes all subjects, `r length(unique(svCompliance$Subject))`, who signed an Informed Consent.  There was no sub-setting based on consent to Genetic or Exploratory Genetic.

Of these subjects, `r length(svAnchorData$PSHSTAT[svAnchorData$PSHSTAT == 'Rollover'])` are rollover subjects.  The rollover status is not entered in the EDC for `r length(svAnchorData$PSHSTAT[is.na(svAnchorData$PSHSTAT)])` subjects.


```{r, echo=FALSE, results = 'asis'}
kable(table(svAnchorData$PSHSTAT, svAnchorData$PSTUDY), 
      caption = 'Direct vs Rollover by Study',
      format = 'html', table.attr = "style='width:30%;'") %>%
  kable_classic(full_width = T, position = 'center',)

```


Some subjects did not complete D1 visit and are not designated as Screen Failure in the EDC.  For this analysis, subjecst without a D1 but with a Study Exit date greater than projected D1 are included in the analysis.
```{r, echo=FALSE}
noSF = svCompliance %>%
  dplyr::filter(is.na(Day_1) & SCRNFAIL_COD == 'N') %>%
  dplyr::select(Subject) %>% 
  unique()

kable(noSF, caption = 'No D1 Non-SF Subjects',
      format = 'html', table.attr = "style='width:20%;'") %>%
  kable_classic(full_width = T, position = 'center',)
```


Reviewing the three laboratory sample data files (FVIII blinded data, un-blinded data and BloodWorks genotyping data) `r length(unique(covLab_raw$SUBJECT))`, `r length(unique(blnd_fviii_raw$SUBJECT))`, `r length(unique(fviiiGeno_raw$Sample.ID))` unique subjects were present, respectively.


During 3 Nov 2020 meeting, CLO PM, Jayson Andrews noted that Blood Works lab will not correct incorrect subject IDs.  The implication is that some lab results reflecting as missing in the original analysis output may be because of incorrect subject IDs.  A comparison was run against Covance and Bloodworks data sets, and the outcome is shown below.
```{r, echo=FALSE}
subCompare1 = covBind %>% dplyr::select(SUBJECT) %>% unique() %>%
  dplyr::anti_join(svAnchorData[, c('Subject')], by = c('SUBJECT' = 'Subject')) %>%
  dplyr::mutate(Source = 'Covance')

subCompare2 = fviiiGeno %>% dplyr::select(SUBJECT) %>% unique() %>%
  dplyr::anti_join(svAnchorData[, c('Subject')], by = c('SUBJECT' = 'Subject')) %>%
  dplyr::mutate(Source = 'Blood Works')

subCompareAll = dplyr::bind_rows(subCompare1, subCompare2)

kable(subCompareAll, caption = 'Subject IDs in Lab Data but not in EDC',
      format = 'html', table.attr = "style='width:80%;'") %>%
  kable_classic(full_width = T, position = 'center',)
      
```
For the two Covance subjects, these are noted to have been screen failed and later re-screened resulting in new subject numbers, 1469-3981 and 0003-39991 respectively.  Covance was requested to update these subject identifiers in their system on 5 Nov 2020.  

Resolution of the Blood Works subjects required linking accession numbers from the incorrect Subject ID to the correct Subject ID.  For example, Subject 1781-3905 samples accession number 6602060219 is linked to subject 1781-3904 in the Covance lab data sets.  A programming fix to account for these discrepancies was applied to the new analysis, but a decision should be make as to whether the programming assumption remains in place or if it must be verified at the vendor first.


### **Visit Data**
The Central Lab Topline Metrics specification document specified test results for Screening and Baseline visits.  Since the protocol allows for Re-screening option, the Smart Re-screening visits were also included in the data set.  The counts of visits included are as follows:


```{r, echo=FALSE}
incVisits = svCompliance %>%
  dplyr::filter(FolderName %in% visitFilter) %>%
  dplyr::filter(Visit_Expected == 'Yes')
  
kable(table(incVisits$OID), caption = 'Expected Visits by OID',
      format = 'html', table.attr = "style='width:30%;'",
      col.names = c("Visit OID", "Freq")) %>%
  kable_classic(full_width = T, position = 'center',)
      
```


### **Lab Data:**

Central Lab Topline Metrics specification document, dated 13Oct2020, requires the following Lab Analyte and corresponding lab sample id's to be used.
```{r, echo=FALSE}
kable(tlmSpecsdf, caption = 'Manual Coding Genotyping Dataset',
      format = 'html', table.attr = "style='width:80%;'") %>%
  kable_classic(full_width = T, position = 'center',)
```


In the intial analysis presented on 30 Oct 2020, 53 FVIII Activity Baseline tests results were reported missing.  Further review of the specs identified duplicate testing types: FVIII Activity Chormogenic vs. FVIII Activity Chormogenic Expedited, and Factor VIII Activity One Stage vs. FVIII Activity OSC Expedited.  When accounting for these, the missing FVIII Activity Baseline test results were reduced significantly.  The table below shows the current mapping to reduce duplicate tests.
```{r, echo=FALSE}
kable(labSkeleton[1:3], caption = 'Manual Coding Genotyping Dataset',
      format = 'html', table.attr = "style='width:80%;'") %>%
  kable_classic(full_width = T, position = 'center',)
```

BMN 270-301_bloodworks_mutation_11Feb2020.csv does not contain LBSPID so these were manually coded.  Further, the lab test names do not match the Data Transfer Specification document so these were also manually coded.  The data set contains multiple tests on each row so a test result was assumed to be present if the "Tested" was present at the beginning of the row.

```{r, echo=FALSE}
kable(fviiiGeno_LBSPID, caption = 'Manual Coding Genotyping Dataset',
      format = 'html', table.attr = "style='width:80%;'") %>%
  kable_classic(full_width = T, position = 'center',)
```

For the remaining two SAS data sets, (bmn_270301_covance_lab_12oct20, bmn_270301_lb_blnd_fviii_12oct20) lab results were defined as missing if the either lab result field was blank or the lab result status was "NOT DONE".

```{r, echo=FALSE}
kable(table(covAll_NoResults$LBSTAT), caption = 'Results Status Options: All Data Sets',
      format = 'html', table.attr = "style='width:80%;'",
      col.names = c("Status", "Freq")) %>%
  kable_classic(full_width = T, position = 'center',)
```


### **Covance Topline Metrics Output**

The updates described above produced considerable changes for the FVIII Activity Baseline category and modest changes for the FVIII Genotype Screening category.  It is anticipated that decisions to adjust the specification document with respect to the HLA Genotpye Screening category may also yield significant decreases in subject's with misisng analyte results.
```{r, echo=FALSE}
kable(covTLM_colapse, caption = 'Subjects Missing One or More Lab Analyte SPID',
      format = 'html', table.attr = "style='width:100%;'") %>%
  kable_classic(full_width = T, position = 'center',)
```



### **Missing Sample Analysis**

Adjustments described in this report show modest to moderate drops in Subject counts except for the HLA Genotype Screening category which reflect the same count as presented in the initial analysis.  

The table below shows the number of missing results by allele.  Allele groupings appears suggesting missing results are not at random.

```{r, echo=FALSE}

covTLM_filter_grp = covTLM_filter %>% dplyr::filter(Covance_Lab_Analyte == 'HLA_Genotype')

ggplot(covTLM_filter_grp,
       aes(x = LBTEST_eqv, fill = 'red')) +
  geom_bar() + 
  geom_text(aes(label = ..count..),
            stat = 'count',
            position = position_stack(vjust = 0.5)) + 
  ggtitle(label = 'Count of Missing Results by Test') +
  labs(x = 'Lab Test',
       y = 'Count Missing Results') +
  theme_light() + 
  scale_x_discrete(labels = function(x) lapply(strwrap(x, 						       
                                                       width = 30,
                                                       simplify = FALSE),
                                               paste, 
                                               collapse = "\n")) +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        legend.position = 'none')

```

Missing results are reviewed by subject and the graph below shows there are 107 subject that are missing all 15 analyte results.  Forty subjects are missing only three of the 15 analytes and nine subjects are missing nine analyte results.
```{r, echo=FALSE}

covTLM_subj_grp = covTLM_filter_grp %>% 
  dplyr::group_by(Subject) %>%
  dplyr::summarise(CountMissing = n()) %>%
  dplyr::ungroup()

ggplot(covTLM_subj_grp,
       aes(x = CountMissing)) + 
  geom_bar(fill = 'light blue') +
  scale_x_discrete(limits = unique(covTLM_subj_grp$CountMissing)) +

  geom_text(aes(label = ..count..),
            stat = 'count',
            position = position_stack(vjust = 0.5)) + 
  ggtitle(label = 'Subject Count by Missing Test Count') +
  
  labs(x = 'Missing Tests',
       y = 'Subject Count') +
  theme_light()

```

The `r length(covTLM_subj_grp$CountMissing[covTLM_subj_grp$CountMissing == 15])` subjects who are missing all 15 lab analyte results are shown below.

```{r, echo=FALSE}
covTLM_subj_grp15 = covTLM_subj_grp %>% filter(CountMissing == 15) %>%
  #dplyr::group_by(Subject) %>%
  dplyr::summarise(SubjectList = paste0(Subject, collapse = ', ')) %>%
  dplyr::select(SubjectList) %>%
  dplyr::ungroup()

knitr::kable(covTLM_subj_grp15,  caption = 'Subjects Missing all Allele Results',
             format = 'html', table.attr = "style='width:100%;'") %>%
  kable_classic(full_width = T, position = 'center',)
```



```{r, echo=FALSE}
covTLM_subj_grp15_ro = covTLM_subj_grp %>% 
  filter(CountMissing == 15) %>%
dplyr::left_join(svAnchorData[, c('Subject', 'PSHSTAT')], by = 'Subject')
```

It was mentioned above that `r length(svAnchorData$PSHSTAT[svAnchorData$PSHSTAT == 'Rollover'])` subjects are rollover. The table above shows that `r length(covTLM_subj_grp15_ro$PSHSTAT[covTLM_subj_grp15_ro$PSHSTAT == 'Rollover'])` of these are rollover.  


Repeating the process for the remaining missing sample groups (9 and 3), subject lists and associated missing tests are shown below.

```{r, echo=FALSE}
covTLM_subj_grp9 = covTLM_subj_grp %>% 
  filter(CountMissing == 9) %>%
  dplyr::summarise(SubjectList = paste0(Subject, collapse = ', ')) %>%
  dplyr::select(SubjectList) %>%
  dplyr::ungroup()

knitr::kable(covTLM_subj_grp9,  caption = 'Subjects Missing Nine Allele Results',
             format = 'html', table.attr = "style='width:100%;'") %>%
  kable_classic(full_width = T, position = 'center',)
```

```{r, echo=FALSE}
covTLM_subj_grp9 = covTLM_subj_grp %>% 
  filter(CountMissing == 9) %>% 
  select(Subject)

covTLM_subj_grp9_test = covTLM_filter_grp %>%
  dplyr::filter(Subject %in% covTLM_subj_grp9$Subject)

ggplot(covTLM_subj_grp9_test,
       aes(x = LBTEST)) + 
  geom_bar(fill = 'light blue') +
  geom_text(aes(label = ..count..),
            stat = 'count',
            position = position_stack(vjust = 0.5)) + 
  ggtitle(label = 'Missing Test Results for Subjects Mising 9 Tests') +
  
  labs(x = 'Missing Test Results',
       y = 'Count') +
  scale_x_discrete(labels = function(x) lapply(strwrap(x, 						       
                                                       width = 20,
                                                       simplify = FALSE),
                                               paste, 
                                               collapse = "\n")) +
  theme_light() + 
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        legend.position = 'none')

```


```{r, echo=FALSE}
covTLM_subj_grp3 = covTLM_subj_grp %>% filter(CountMissing == 3) %>%
  #dplyr::group_by(Subject) %>%
  dplyr::summarise(SubjectList = paste0(Subject, collapse = ', ')) %>%
  dplyr::select(SubjectList) %>%
  dplyr::ungroup()

knitr::kable(covTLM_subj_grp3,  caption = 'Subjects Missing Nine Allele Results',
             format = 'html', table.attr = "style='width:100%;'") %>%
  kable_classic(full_width = T, position = 'center',)
```


```{r, echo=FALSE}
covTLM_subj_grp3 = covTLM_subj_grp %>% 
  filter(CountMissing == 3) %>% 
  select(Subject)

covTLM_subj_grp3_test = covTLM_filter_grp %>%
  dplyr::filter(Subject %in% covTLM_subj_grp3$Subject)

ggplot(covTLM_subj_grp3_test,
       aes(x = LBTEST)) + 
  geom_bar(fill = 'light blue') +
  geom_text(aes(label = ..count..),
            stat = 'count',
            position = position_stack(vjust = 0.5)) + 
  ggtitle(label = 'Missing Test Results for Subjects Mising 3 Tests') +
  
  labs(x = 'Missing Test Results',
       y = 'Count') +
  theme_light() + 
  scale_x_discrete(labels = function(x) lapply(strwrap(x, 						       
                                                       width = 20,
                                                       simplify = FALSE),
                                               paste, 
                                               collapse = "\n")) +
  theme(axis.text.x = element_text(angle = 45,
                                   hjust = 1),
        legend.position = 'none')

```


## **Conclusion**

Even with adjustments to the original code, the HLA group contained a high number of missing test results.  Anlaysis of these missing tests suggests they are not missing and random and additional SME input is needed to be able to appropriately spec this listing and update programming to account for HLA tests that are not expected.



